
# Measurements along the way / libs / notes etc.
```{r}

# par(mfrow=c(2,2))
# hist(df01$population_01042017, breaks = 80)
# #png(filename= "hist_pop_norm.png",height = 600, width = 600)
# hist(df01$pop_norm, breaks = 80)
# #png(filename= "hist_q_pop_real.png",height = 600, width = 600)
# hist(quantile(df01$population_01042017), breaks = 80)
# #png(filename= "hist_q_pop_norm.png",height = 600, width = 600)
# hist(quantile(df01$pop_norm), breaks = 80)
# png(filename= "hists_pop.png", width =800,height=800)
# dev.off()
# 
# # hst1 <- hist(df01$population_01042017, breaks =   120, xaxt="n")
# # axis(side=1, at=axTicks(1), labels=formatC(axTicks(1), format="d", digits = 11, big.mark=','))
# hst2 <- hist(df01$pop_norm, breaks = 120, xaxt="n")
# axis(side=1, at=axTicks(1), labels=formatC(axTicks(1), format="d", digits = 11, big.mark=','))


# > nrow(subset(df01, df01$population_01042017 > 100000))
# [1] 31
# > head(subset(df01, df01$population_01042017 > 100000))
#           municipality population_01042017 longitude latitude  pop_norm
# 8              Alkmaar              108618  4.753375 52.63238 0.1263213
# 10              Almere              201703  5.264702 52.35078 0.2355291
# 11 Alphen aan den Rijn              109107  4.668851 52.12766 0.1268950
# 14          Amersfoort              154712  5.387827 52.15611 0.1803990
# 16           Amsterdam              853312  4.895168 52.37022 1.0000000
# 17           Apeldoorn              160497  5.969923 52.21116 0.1871860
# 
# > nrow(subset(df01, df01$population_01042017 > 200000))
# [1] 8
# > head(subset(df01, df01$population_01042017 > 200000))
#                 municipality population_01042017 longitude latitude  pop_norm
# 10                    Almere              201703  5.264702 52.35078 0.2355291
# 16                 Amsterdam              853312  4.895168 52.37022 1.0000000
# 95                 Eindhoven              227100  5.469722 51.44164 0.2653250
# 120 's-Gravenhage (gemeente)              526439  4.300700 52.07050 0.6165110
# 121     Groningen (gemeente)              202250  6.551335 53.23239 0.2361708
# 281                Rotterdam              639587  4.477733 51.92442 0.7492568
# 
# > nrow(subset(df01, df01$population_01042017 > 500000))
# [1] 3
# > head(subset(df01, df01$population_01042017 > 500000))
#                 municipality population_01042017 longitude latitude  pop_norm
# 16                 Amsterdam              853312  4.895168 52.37022 1.0000000
# 120 's-Gravenhage (gemeente)              526439  4.300700 52.07050 0.6165110
# 281                Rotterdam              639587  4.477733 51.92442 0.7492568


# quantile(df01$pop_population_01042017)
#          0%       25%        50%      75%     100% 
#      946.00  17847.25  26911.50  44508.25 853312.00 
#
# quantile(df01$pop_norm)
#         0%        25%        50%        75%       100% 
# 0.00000000 0.01982863 0.03046285 0.05110745 1.00000000 


#install.packages(“ggplot2″)
#install.packages(“jsonlite”)
#install.packages(“plyr”)
#install_github("leonawicz/mapmate")
#install_github("wjrl/RBioFabric")

# df03_possible_duplicates <- duplicated(df03$distances)
# df03_uniqdists <- unique(df03$distances)
#
# df03$possible_duplicate <- 2
# # remove duplicate connections
# for (i in 1:nrow(df03)){
#   for (j in 1:800){
#     ifelse(
#       (df03$distances[i] == df03$distances[j]),
#         df03$possible_duplicate[i] <- 1,
#         df03$possible_duplicate[i] <- 0
#     )}}
# df03_possible_duplicates <- duplicated(df03[,5:6])
# nrow(df03$possible_duplicate)
 

```
 
# 0 - Start: set wd, read in data, load libs -> df01
```{r}

#setwd("/x/y/z/")

library(RJSONIO)
library(ggmap)
library(geosphere)
library(data.table)
library(gdata)
library(varhandle)
library(mapmate)
library(dplyr)
library(igraph)
library(network)
library(sna)
library(ndtv)
library(visNetwork)
library(maps)
library(ggmap)
library(mapdata)
library(mapproj)
library(maptools)
library(RColorBrewer)
library(classInt)
#library(rgdal)
library(scales)
library(raster)
library(RBioFabric)
library(HiveR)

df00 <- read.csv("Bevolkingsontwikkeli_040617192441.csv",header = TRUE)

# get only the location-name and population variables
df01 <- df00[1:388,c(1,17)]

# get lon & lat values from google maps api from each municipality based on location names
#geocodes <- geocode(as.character(df01$Regio.s))
#write.csv(geocodes,"geocodes.csv")
geo <- read.csv("geocodes.csv",header=TRUE)
df01 <- data.frame(df01[,1:2],geo)
df01 <- df01[,-c(3)]

```

# 1 - First renaming and processing (normalizing population, adding 3 city-size variables)
```{r}

colnames(df01)[1] <- "municipality"
colnames(df01)[2] <- "population_01042017"
colnames(df01)[3] <- "longitude"
colnames(df01)[4] <- "latitude"

df01$population_01042017 <- as.integer(df01$population_01042017)

# easier for comparison, and slightly anonimized
df01$pop_round_100 <- round(df01$population_01042017 / 100)
df01$pop_norm <- ( (df01$pop_round_100 - min(df01$pop_round_100))
                   / (max(df01$pop_round_100) - min(df01$pop_round_100)) )

# all 3 df df01$ [pop_10e5_20e5, pop_over_20e5,pop_under_10e5] 
    df01$pop_10e5_20e5 <- 20
    for (i in 1:nrow(df01)){
    ifelse(
      ((df01$pop_round_100[i] >= as.integer(1000)) && (df01$pop_round_100 < as.integer(2000))) 
      , df01$pop_10e5_20e5[i] <- 1  
      , ifelse(
        ((df01$pop_round_100[i] >= as.integer(2000)) || (df01$pop_round_100 < as.integer(1000)))
        , df01$pop_10e5_20e5[i] <- 0
        , df01$pop_10e5_20e5[i] <- 1)
    )}
    df01$pop_over_20e5 <- 20
    for (i in 1:nrow(df01)){
    ifelse(
      (df01$pop_round_100[i] >= as.integer(2000))
      , df01$pop_over_20e5[i] <- 1
      , ifelse(
        (df01$pop_round_100[i] < as.integer(2000))
        , df01$pop_over_20e5[i] <- 0
        , df01$pop_over_20e5[i] <- 1)
    )} 
    df01$pop_under_10e5 <- 20
    for (i in 1:nrow(df01)){
    ifelse(
      ((as.integer(df01$pop_10e5_20e5[i]) == as.integer(0))
       && (as.integer(df01$pop_over_20e5[i]) == as.integer(0)))
      , df01$pop_under_10e5[i] <- 1
      , ifelse(
        ((as.integer(df01$pop_10e5_20e5[i]) == as.integer(1))
         || (as.integer(df01$pop_over_20e5[i]) == as.integer(1)))
        , df01$pop_under_10e5[i] <- 0
        , df01$pop_under_10e5[i] <- 1)
    )}

   df01$pop_norm_0102 <- 20
    for (i in 1:nrow(df01)){
    ifelse(
      ((as.numeric(df01$pop_norm)[i] >= as.numeric(0.1)) && (as.numeric(df01$pop_norm) < as.numeric(0.2))) 
      , df01$pop_norm_0102[i] <- 1  
      , ifelse(
        ((as.numeric(df01$pop_norm)[i] >= as.numeric(0.2)) || (as.numeric(df01$pop_norm) < as.numeric(0.1)))
        , df01$pop_norm_0102[i] <- 0
        , df01$pop_norm_0102[i] <- 1)
    )}
    df01$pop_norm_02up <- 20
    for (i in 1:nrow(df01)){
    ifelse(
      (as.numeric(df01$pop_norm)[i] >= as.numeric(0.2))
      , df01$pop_norm_02up[i] <- 1
      , ifelse(
        (as.numeric(df01$pop_norm)[i] < as.numeric(0.2))
        , df01$pop_norm_02up[i] <- 0
        , df01$pop_norm_02up[i] <- 1)
    )} 
    df01$pop_norm_below01 <- 20
    for (i in 1:nrow(df01)){
    ifelse(
      ((as.integer(df01$pop_norm_0102[i]) == as.integer(0))
       && (as.integer(df01$pop_norm_02up[i]) == as.integer(0)))
      , df01$pop_norm_below01[i] <- 1
      , ifelse(
        ((as.integer(df01$pop_norm_0102[i]) == as.integer(1))
         || (as.integer(df01$pop_norm_02up[i]) == as.integer(1)))
        , df01$pop_norm_below01[i] <- 0
        , df01$pop_norm_below01[i] <- 1)
    )}
 
```

# 2 - Matrix, removal of double 'edges', transposing, etc
```{r}

# create a distance matrix from two lists (-1 on both first and last city-row)
l1 <- data.frame(longitude = df01[1:387,3],
                 latitude = df01[1:387,4])
l2 <- data.frame(longitude = df01[2:388,3],
                 latitude = df01[2:388,4])
mtx <- distm(l1[,c('longitude','latitude')], l2[,c('longitude','latitude')], fun=distVincentyEllipsoid)

#dim(mtx)
# [1] 387 387
# elements in mtx = 149769
# checked: 387*387 = 149769

# remove duplicates and 'eigen'-distances
mtx[lower.tri(mtx)] <- NA
#head(mtx)
#         [,1]      [,2]     [,3]      [,4]      [,5]      [,6]
# [1,] 181411 158998.75 121147.8  45689.08 189958.68 204965.68
# [2,]     NA  60463.11 106807.8 178628.27  29788.73  46693.60
# [3,]     NA        NA 130832.1 141961.74  43773.50  51019.39
# [4,]     NA        NA       NA 145751.95 131649.72 150232.45
#dim(mtx) # > [1] 387 387

# transpose as quick-fix for correct iteration direction in as.list
mx <- t(mtx)
dists <- as.list(mx)
dists2 <- dists[!is.na(dists)]
dists3 <- data.frame(dists2)
dists4 <- t(dists3)
dists <- data.frame(dists4)

# ROUND_DISTANCES will be handy for further processing
# it introcuses a (tolerable?) error of <50 cm. (w/in one distance-measurement!)
dists$dists_km <- (dists$dists4 / 1000)
round_distances <- as.integer(round(dists$dists_km))
dists_norm <- ( (round_distances-min(round_distances))/(max(round_distances)-min(round_distances)) )

# dim(distances) # > [1] 75078     1 # checked: 149769 + 387 / 2 = 75078

```

# 3a - Get all 'pairwise' combinations; combine all variables
#    -- MERGING
```{r}
# dataframe should (row-wise) increase from 388 to 75078 (see above)
# get X-combinations per Y w/ 'combn' and split into seperate variables

# municipalities
compared_municipalities <- data.frame(combn(as.character(df01$municipality), 2, FUN = paste, collapse="_"))
# locations 
combined_lonlats <- paste(df01$longitude, df01$latitude, sep = ",")
compared_locations <- data.frame(combn(as.character(combined_lonlats), 2, FUN = paste, collapse="_"))
# populations
compared_pop_round <- data.frame(combn(as.character(df01$pop_round_100), 2, FUN = paste, collapse="_"))
compared_pop_norm <- data.frame(combn(as.character(df01$pop_norm), 2, FUN = paste, collapse="_"))
compared_pop_under_10e5 <- data.frame(combn(as.character(df01$pop_under_10e5), 2, FUN = paste, collapse="_"))
compared_pop_10e5_20e5 <- data.frame(combn(as.character(df01$pop_10e5_20e5), 2, FUN = paste, collapse="_"))
compared_pop_over_20e5 <- data.frame(combn(as.character(df01$pop_over_20e5), 2, FUN = paste, collapse="_"))
compared_pop_norm_below01 <- data.frame(combn(as.character(df01$pop_norm_below01), 2, FUN = paste, collapse="_")) 
compared_pop_norm_0102 <- data.frame(combn(as.character(df01$pop_norm_0102), 2, FUN = paste, collapse="_")) 
compared_pop_norm_02up <- data.frame(combn(as.character(df01$pop_norm_02up), 2, FUN = paste, collapse="_")) 

df012 <- data.frame(compared_municipalities, compared_locations)
df013 <- data.frame(df012, dists)
df014 <- data.frame(df013, dists_norm)
df015 <- data.frame(df014, compared_pop_round)
df016 <- data.frame(df015, compared_pop_norm)
df017 <- data.frame(df016, compared_pop_under_10e5)
df018 <- data.frame(df017, compared_pop_10e5_20e5)
df019 <- data.frame(df018, compared_pop_over_20e5)
df020 <- data.frame(df019, compared_pop_norm_below01)
df021 <- data.frame(df020, compared_pop_norm_0102)
df02 <- data.frame(df021, compared_pop_norm_02up)
  
df02 <- df02[,-3]

# rename column names
colnames(df02)[1] <- "municipalities_compared"
colnames(df02)[2] <- "locations_compared"
colnames(df02)[3] <- "distances"
colnames(df02)[4] <- "dists_normalized"
colnames(df02)[5] <- "pop_round_compared"
colnames(df02)[6] <- "pop_norm_compared"
colnames(df02)[7] <- "pop_under_10e5_compared"
colnames(df02)[8] <- "pop_10e5_20e5_compared"
colnames(df02)[9] <- "pop_over_20e5_compared"
colnames(df02)[10] <- "pop_norm_below01_compared"
colnames(df02)[11] <- "pop_norm_0102_compared"
colnames(df02)[12] <- "pop_norm_02up_compared"

```

#    -- SPLITTING
```{r}
# make seperate lists with split-up's of compared variables

# munis
lcm <- data.frame(strsplit(as.character(df02$municipalities_compared), '_'))
lcm2 <- data.frame(t(lcm))
# locs
lcl <- data.frame(strsplit(as.character(df02$locations_compared), '_'))
lcl2 <- data.frame(t(lcl))
# pops
lcp <- data.frame(strsplit(as.character(df02$pop_round_compared), '_'))
lcp2 <- data.frame(t(lcp))
lcpn <- data.frame(strsplit(as.character(df02$pop_norm_compared), '_'))
lcpn2 <- data.frame(t(lcpn))
lcppu1 <- data.frame(strsplit(as.character(df02$pop_under_10e5_compared), '_'))
lcppu12 <- data.frame(t(lcppu1))
lcp12 <- data.frame(strsplit(as.character(df02$pop_10e5_20e5_compared), '_'))
lcp122 <- data.frame(t(lcp12))
lcpo2 <- data.frame(strsplit(as.character(df02$pop_over_20e5_compared), '_'))
lcpo22 <- data.frame(t(lcpo2))
lcpnb1 <- data.frame(strsplit(as.character(df02$pop_norm_below01_compared), '_'))
lcpnb12 <- data.frame(t(lcpnb1))
lcpn12 <- data.frame(strsplit(as.character(df02$pop_norm_0102_compared), '_'))
lcpn122 <- data.frame(t(lcpn12))
lcpn2u <- data.frame(strsplit(as.character(df02$pop_norm_02up_compared), '_'))
lcpn2u2 <- data.frame(t(lcpn2u))
# merge back into one dataframe
df020 <- data.frame(df02, lcm2)
df021 <- data.frame(df020, lcl2)
df022 <- data.frame(df021, lcp2)
df023 <- data.frame(df022, lcpn2)
df024 <- data.frame(df023, lcppu12)
df025 <- data.frame(df024, lcp122)
df026 <- data.frame(df025, lcpo22)
df027 <- data.frame(df026, lcpnb12)
df028 <- data.frame(df027, lcpn122)
df03 <- data.frame(df028, lcpn2u2)

colnames(df03)[13] <- "muni_i"
colnames(df03)[14] <- "muni_j"
colnames(df03)[15] <- "lonlat_i"
colnames(df03)[16] <- "lonlat_j"
colnames(df03)[17] <- "pop_i"
colnames(df03)[18] <- "pop_j"
colnames(df03)[19] <- "pop_norm_i"
colnames(df03)[20] <- "pop_norm_j"
colnames(df03)[21] <- "pop_under_10e5_i" 
colnames(df03)[22] <- "pop_under_10e5_j"
colnames(df03)[23] <- "pop_10e5_20e5_i"
colnames(df03)[24] <- "pop_10e5_20e5_j"
colnames(df03)[25] <- "pop_over_20e5_i"
colnames(df03)[26] <- "pop_over_20e5_j"
colnames(df03)[27] <- "pop_norm_below01_i"
colnames(df03)[28] <- "pop_norm_below01_j"
colnames(df03)[29] <- "pop_norm_0102_i"
colnames(df03)[30] <- "pop_norm_0102_j"
colnames(df03)[31] <- "pop_norm_02up_i"
colnames(df03)[32] <- "pop_norm_02up_j"

lcl_lli <- data.frame(strsplit(as.character(df03$lonlat_i), ','))
lcl_lli2 <- data.frame(t(lcl_lli), stringsAsFactors = TRUE)
lcl_llj <- data.frame(strsplit(as.character(df03$lonlat_j), ','))
lcl_llj2 <- data.frame(t(lcl_llj), stringsAsFactors = TRUE)
 
df033 <- data.frame(lcl_lli2, lcl_llj2)
df04 <- data.frame(df03, df033)

colnames(df04)[33] <- "lon_i"
colnames(df04)[34] <- "lat_i"
colnames(df04)[35] <- "lon_j"
colnames(df04)[36] <- "lat_j"

df05 <- df04[,-c(1,2,5,6,7,8,9,10,11,12,15,16)]

```

# 4 - Changing types etc.
```{r}
# munis, pops, etc.
a <- 3:4
df05[a] <- lapply(df05[a], as.character)
b <- 7:8
df05[b] <- lapply(df05[b], unfactor)
c <- 9:20
df05[c] <- lapply(df05[c], unfactor)
df05[c] <- lapply(df05[c], as.integer)
d <- 21:24
df05[d] <- lapply(df05[d], unfactor)

# backup!
write.csv(df05, 'df05.csv')

```

# Directionality & movement
```{r}
# for drawing directed edges later on
# decide DIRECTION based on the larger population out of any 2-combination
# the hypothesis is that the smaller population will be 'drawn' to the larger population

df05$from_lon <- 10
for (i in 1:nrow(df05)){
  ifelse(
    (df05$pop_norm_i[i] < df05$pop_norm_j[i])
    , df05$from_lon[i] <- df05$lon_i[i]
      , ifelse(
        (df05$pop_norm_i[i] > df05$pop_norm_j[i])
        , df05$from_lon[i] <- df05$lon_j[i]
        , df05$from_lon[i] <- 0)
  )}

df05$from_lat <- 10
for (i in 1:nrow(df05)){
  ifelse(
    (df05$pop_norm_i[i] < df05$pop_norm_j[i])
    , df05$from_lat[i] <- df05 $lat_i[i]
      , ifelse(
        (df05$pop_norm_i[i] > df05$pop_norm_j[i])
        , df05$from_lat[i] <- df05$lat_j[i]
        , df05$from_lat[i] <- 0)
  )}

df05$to_lon <- 10
for (i in 1:nrow(df05)){
  ifelse(
    (df05$pop_norm_i[i] > df05$pop_norm_j[i])
    , df05$to_lon[i] <- df05$lon_i[i]
    , ifelse(
      (df05$pop_norm_i[i] < df05$pop_norm_j[i])
      , df05$to_lon[i] <- df05$lon_j[i]
      , df05$to_lon[i] <- 0)
  )}

df05$to_lat <- 10
for (i in 1:nrow(df05)){
  ifelse(
    (df05$pop_norm_i[i] > df05$pop_norm_j[i])
    , df05$to_lat[i] <- df05 $lat_i[i]
    , ifelse(
      (df05$pop_norm_i[i] < df05$pop_norm_j[i])
      , df05$to_lat[i] <- df05$lat_j[i]
      , df05$to_lat[i] <- 0)
  )}

# chance of actual movement, as being not 'too far' (too simplistic approach!)
# draw edge at each distance over threshold; thresholds for now are 50000 and 100000 (m)
# also, actual trip-duraion should be taken into acount, regardles of (spherical) lon-lat distances
# note: 2 more gradients should be added for more realistic choices

df05$edge_dist100km_up <- 2
for (i in 1:nrow(df05))
{
  ifelse(
    (df05$distances[i] >= as.numeric(100))
    , df05$edge_dist100km_up[i] <- 1
    , ifelse(
      (df05$distances[i] < as.numeric(100))
      , df05$edge_dist100km_up[i] <- 0
      , df05$edge_dist100km_up[i] <- 1)
  )}

df05$edge_dist50_100km <- 2
for (i in 1:nrow(df05))
{
  ifelse(
  ((df05$distances[i] >= as.numeric(50)) && (df05$distances < as.numeric(100))) 
    , df05$edge_dist50_100km[i] <- 1
    , ifelse(
     ((df05$distances[i] <= as.numeric(50)) || (df05$distances > as.numeric(100))) 
      , df05$edge_dist50_100km[i] <- 0
      , df05$edge_dist50_100km[i] <- 1)
  )}

df05$edge_dist_below50km <- 2
for (i in 1:nrow(df03)){
  ifelse(
  (df05$distances[i] <= as.numeric(50)) 
    , df05$edge_dist_below50km[i] <- 1
    , ifelse(
     (df05$distances[i] >= as.numeric(50)) 
      , df05$edge_dist_below50km[i] <- 0
      , df05$edge_dist_below50km[i] <- 1)
  )}

# the gravity is based on the 'grivity model' ((pop_i * pop_j) / distances)
df05$gravity2 <- 0
for (i in 1:nrow(df05)){
    df05$gravity2[i] <- 
                        (round
                           (
                             (
                             as.integer(round(df05$pop_norm_i[i]*1000))
                             *
                             as.integer(round(df05$pop_norm_j[i]*1000))
                             ) 
                             /
                             as.integer(round(df05$dists_normalized[i]*100), digits = 2)
                             , digits = 2
                           )
                         )
                      }

df05$gravity_norm <- ( (df05$gravity - min(df05$gravity)) / (max(df05$gravity) - min(df05$gravity)) )

write.csv(df05, 'df05_w_edges_grav.csv')

```

# TMP START0
```{r}
#setwd("/x/y/z/")

library(RJSONIO)
library(ggmap)
library(geosphere)
library(data.table)
library(gdata)
library(varhandle)
library(mapmate)
library(dplyr)
library(igraph)
library(network)
library(sna)
library(ndtv)
library(threejs)
library(networkD3)
library(visNetwork)
library(maps)
library(ggmap)
library(mapdata)
library(mapproj)
library(maptools)
library(RColorBrewer)
library(classInt)
#library(rgdal)
library(scales)
library(raster)
library(RBioFabric)
library(HiveR)


df05 <- read.csv("df05_w_edges_grav.csv",header = TRUE)

```

# Subsetting for plots
```{r}

# subsetting only necessary values, new dataframes for network & map plots 
# filter only 'actual' (expected, by naif estimations) movements

# df_map_plot <- data.frame(df04[,c(6,7,16,17,18,19,21,22,23,24,25)])

df05$gravity <- as.integer(df05$gravity)
df05$gravity_norm<- as.integer(df05$gr)
df_map_plot <- subset(df05, edge_dist50km == 1)
#df_map_plot <- subset(df_map_plot[, c(1,2,3,4)])

df_map$gravity <- as.integer(df_map$gravity)
df_map$gravity <- as.integer(df_map$gravity)
df_map_s2 <- subset(df_map, normalized_gravity > 0.01)
df_map <- df_map_plot 
df_map$gravity <- as.integer(df_map$gravity)
df_map$normalized_gravity <- (df_map$gravity-min(df_map$gravity))/(max(df_map$gravity)-min(df_map$gravity))
df_map_s2 <- subset(df_map, normalized_gravity > 0.02)

# df_map_s2 <- subset(df_map, normalized_gravity > 0.01)
# > dim(df_map_s2)
# [1] 1959    9

df_network_plot <- data.frame(df04[,c(6,7,21,23)])
df_network_plot <- subset(df_network_plot, edge2 == 1)
df_network_plot$gravity <- as.integer(df_network_plot$gravity)
df_network_plot$normalized_gravity <- (df_network_plot$gravity-min(df_network_plot$gravity))/(max(df_network_plot$gravity)-min(df_network_plot$gravity))
df_network_plot2<- subset(df_network_plot, normalized_gravity > 0.1)
df_network_plot3 <- df_network_plot2[,c(1,2,5)]



```
 
