

 
# Start: read in data, load libs -> df01
```{r}
 
# set wd, load libs, read initial csv into dataframe
setwd("/home/bigdata09/projs/mob/")

df00 <- read.csv("Bevolkingsontwikkeli_040617192441.csv",header = TRUE)
#dim(df00)
# [1] 388  17

# get only the location-name and population variables
df01 <- df00[1:388,c(1,17)]

# get lon & lat values from google maps api from each municipality based on location names
#geocodes <- geocode(as.character(df01$Regio.s))
#write.csv(geocodes,"geocodes.csv")
geo <- read.csv("geocodes.csv",header=TRUE)
df01 <- data.frame(df01[,1:2],geo)
df01 <- df01[,-c(3)]

# > min(df01[,2])
# [1] 946
# > min(df01[,2])
# [1] 946

```

# First renaming and processing (normalizing population, adding 3 city-size variables)

```{r}

colnames(df01)[1] <- "municipality"
colnames(df01)[2] <- "population_01042017"
colnames(df01)[3] <- "longitude"
colnames(df01)[4] <- "latitude"

df01$population_01042017 <- as.integer(df01$population_01042017)

df01$pop_01 <- ( (df01$population_01042017 - min(df01$population_01042017)) 
               / (max (df01$population_01042017) - min(df01$population_01042017)) )

# hst1 <- hist(df01$population_01042017, breaks = 120, xaxt="n")
# axis(side=1, at=axTicks(1), labels=formatC(axTicks(1), format="d", digits = 11, big.mark=','))
# hst2 <- hist(df01$pop_01, breaks = 20, xaxt="n")
# axis(side=1, at=axTicks(1), labels=formatC(axTicks(1), format="d", digits = 11, big.mark=','))

df01$pop_10e5_20e5 <- 20
for (i in 1:nrow(df01)){
  ifelse(
    ((df01$population_01042017[i] >= as.integer(100000)) && (df01$population_01042017 < as.integer(200000))) 
    , df01$pop_10e5_20e5[i] <- 1  
    , ifelse(
      ((df01$population_01042017[i] >= as.integer(200000)) || (df01$population_01042017 < as.integer(100000)))
      , df01$pop_10e5_20e5[i] <- 0
      , df01$pop_10e5_20e5[i] <- 1)
  )
}

df01$pop_over_20e5 <- 20
for (i in 1:nrow(df01)){
  ifelse(
    (df01$population_01042017[i] >= as.integer(200000))
    , df01$pop_over_20e5[i] <- 1
    , ifelse(
      (df01$population_01042017[i] < as.integer(200000))
      , df01$pop_over_20e5[i] <- 0
      , df01$pop_over_20e5[i] <- 1)
  )
} 

df01$pop_under_10e5 <- 20
for (i in 1:nrow(df01)){
  ifelse(
    ((as.integer(df01$pop_10e5_20e5_i[i]) == as.integer(0))
     && (as.integer(df01$pop_10e5_20e5_j[i]) == as.integer(0))
     && (as.integer(df01$pop_over_20e5_i[i]) == as.integer(0))
     && (as.integer(df01$pop_over_20e5_j[i]) == as.integer(0)))
    , df01$pop_under_10e5[i] <- 1
    , ifelse(
      ((as.integer(df01$pop_10e5_20e5_i[i]) == as.integer(1))
       || (as.integer(df01$pop_10e5_20e5_j[i]) == as.integer(1))
       || (as.integer(df01$pop_over_20e5_i[i]) == as.integer(1))
       || (as.integer(df01$pop_over_20e5_j[i]) == as.integer(1)))
      , df01$pop_under_10e5[i] <- 0
      , df01$pop_under_10e5[i] <- 1)
  )
}

#dim(df01)
# [1] 388   7
 
```


# Matrix, removal of double 'edges', transposing, etc

```{r}

# create a distance matrix from two lists (-1 on both first and last city-row)
l1 <- data.frame(longitude = df01[1:387,3],
                 latitude = df01[1:387,4])
l2 <- data.frame(longitude = df01[2:388,3],
                 latitude = df01[2:388,4])
mtx <- distm(l1[,c('longitude','latitude')], l2[,c('longitude','latitude')], fun=distVincentyEllipsoid)

#dim(mtx)
# [1] 387 387
# elements in mtx = 149769
# checked: 387*387 = 149769

# remove duplicates and 'eigen'-distances
mtx[lower.tri(mtx)] <- NA
#head(mtx)
#         [,1]      [,2]     [,3]      [,4]      [,5]      [,6]
# [1,] 181411 158998.75 121147.8  45689.08 189958.68 204965.68
# [2,]     NA  60463.11 106807.8 178628.27  29788.73  46693.60
# [3,]     NA        NA 130832.1 141961.74  43773.50  51019.39
# [4,]     NA        NA       NA 145751.95 131649.72 150232.45
#dim(mtx)
# [1] 387 387

# transpose as quick-fix for correct iteration direction in as.list
mx <- t(mtx)
dists <- as.list(mx)
dists2 <- dists[!is.na(dists)]
dists3 <- data.frame(dists2)
dists4 <- t(dists3)
distances <- data.frame(dists4)

# dim(distances)
# [1] 75078     1
# checked: 149769 + 387 / 2 = 75078

```

# Get all 'pairwise' combinations; combine all variables and split for further comparison

```{r}
# dataframe should (row-wise) increase from 388 to 75078 (see above)
# get X-combinations per Y w/ 'combn' and split into seperate variables

# municipalities
compared_municipalities <- data.frame(combn(as.character(df01$municipality), 2, FUN = paste, collapse="_"))

# populations
compared_populations <- data.frame(combn(as.character(df01$population_01042017), 2, FUN = paste, collapse="_"))
compared_pop_01 <- data.frame(combn(as.character(df01$pop_01), 2, FUN = paste, collapse="_"))

# locations 
combined_lonlats <- paste(df01$longitude, df01$latitude, sep = ",")
compared_locations <- data.frame(combn(as.character(combined_lonlats), 2, FUN = paste, collapse="_"))

# (big) cities
comb_pop_under_10e5 <- data.frame(combn(as.integer(df01$pop_under_10e5), 2, FUN = paste, collapse="_"))
comb_pop_10e5_20e5 <- data.frame(combn(as.integer(df01$pop_10e5_20e5), 2, FUN = paste, collapse="_"))
comb_pop_over_20e5 <- data.frame(combn(as.character(df01$pop_over_20e5), 2, FUN = paste, collapse="_"))

# into a new dataframe, should be 75078x6
df011 <- data.frame(compared_municipalities, distances)
df012 <- data.frame(df011, compared_populations)
df013 <- data.frame(df012, compared_pop_01)
df014 <- data.frame(df013, compared_locations)
df015 <- data.frame(df014, pop_under_10e5)
df016 <- data.frame(df015, comb_pop_10e5_20e5)
df02 <- data.frame(df016, comb_pop_over_20e5)

# rename column names
colnames(df02)[1] <- "compared_municipalities"
colnames(df02)[2] <- "distances"
colnames(df02)[3] <- "compared_populations"
colnames(df02)[4] <- "compared_pop_01"
colnames(df02)[5] <- "compared_locations"
colnames(df02)[6] <- "comb_pop_under_10e5"
colnames(df02)[7] <- "comb_pop_10e5_20e5"
colnames(df02)[8] <- "comb_pop_over_20e5"


# make seperate lists with split-up's of compared variables

# munis
lcm <- data.frame(strsplit(as.character(df02$compared_municipalities), '_'))
lcm2 <- data.frame(t(lcm))

# pops
lcp <- data.frame(strsplit(as.character(df02$compared_populations), '_'))
lcp2 <- data.frame(t(lcp))
lcp01 <- data.frame(strsplit(as.character(df02$compared_pop_01), '_'))
lcp012 <- data.frame(t(lcp01))

# locs
lcl <- data.frame(strsplit(as.character(df02$compared_locations), '_'))
lcl2 <- data.frame(t(lcl))

# pop scales
lccunder10 <- data.frame(strsplit(as.character(df02$comb_pop_under_10e5), '_'))
lccunder102 <- data.frame(t(lccunder10))
lcc1020 <- data.frame(strsplit(as.character(df02$comb_pop_10e5_20e5), '_'))
lcc10202 <- data.frame(t(lcbc1020))
lcc20up <- data.frame(strsplit(as.character(df02$comb_pop_over_20e5), '_'))
lcc20up2 <- data.frame(strsplit(t(lcbc20up)))
  
# merge back into one dataframe
df020 <- data.frame(df02, lcm2)
df021 <- data.frame(df020, lcp2)
df022 <- data.frame(df021, lcp012)
df023 <- data.frame(df022, lcl2)
df024 <- data.frame(df023, lccunder102)
df025 <- data.frame(df024, lcc10202)
df026 <- data.frame(df025, lcc20up2)

colnames(df026)[1] <- "compared_municipalities"
colnames(df026)[2] <- "distances"
colnames(df026)[3] <- "compared_populations"
colnames(df026)[4] <- "compared_normal_populations"
colnames(df026)[5] <- "compared_locations"
colnames(df026)[6] <- "muni_i"
colnames(df026)[7] <- "muni_j"
colnames(df026)[8] <- "pop_i"
colnames(df026)[9] <- "pop_j"
colnames(df026)[10] <- "lonlat_i"
colnames(df026)[11] <- "lonlat_j"
colnames(df026)[12] <- "comb_pop_under_10e5"
colnames(df026)[13] <- "comb_pop_10e5_20e5"
colnames(df026)[14] <- "comb_pop_over_20e5"

df027 <- data.frame(lcl_lons2, lcl_lats2)
df03 <- data.frame(df026, df025)
  
colnames(df03)[13] <- "lon_i"
colnames(df03)[14] <- "lat_i"
colnames(df03)[15] <- "lon_j"
colnames(df03)[16] <- "lat_j"

```

```{Plots}

plot(z)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


```{backups etc}



#install.packages(“ggplot2″)
#install.packages(“jsonlite”)
#install.packages(“plyr”)
#install_github("leonawicz/mapmate")
#install_github("wjrl/RBioFabric")



```
