
# 1 - First renaming and processing (normalizing population, adding 3 city-size variables)
```{r}

colnames(df01)[1] <- "municipality"
colnames(df01)[2] <- "population_01042017"
colnames(df01)[3] <- "longitude"
colnames(df01)[4] <- "latitude"

df01$population_01042017 <- as.integer(df01$population_01042017)

# easier for comparison, and slightly anonimized
df01$pop_round_100 <- round(df01$population_01042017 / 100)
pop_round_100 <- round(df01$population_01042017 / 100)
df01$pop_z <- scale(df01$pop_round_100, center = TRUE, scale = TRUE)

# all 3 df df01$ [pop_10e5_20e5, pop_over_20e5,pop_under_10e5] 
    df01$pop_z_1020 <- 20
    for (i in 1:nrow(df01)){
    ifelse(
      ((df01$pop_z[i] >= as.integer(1)) && (df01$pop_z < as.integer(2))) 
      , df01$pop_z_1020[i] <- 1  
      , ifelse(
        ((df01$pop_z[i] >= as.integer(2)) || (df01$pop_z < as.integer(1)))
        , df01$pop_z_1020[i] <- 0
        , df01$pop_z_1020[i] <- 1)
    )}
    df01$pop_z_20u <- 20
    for (i in 1:nrow(df01)){
    ifelse(
      (df01$pop_z[i] >= as.integer(2))
      , df01$pop_z_20u[i] <- 1
      , ifelse(
        (df01$pop_z[i] < as.integer(2))
        , df01$pop_z_20u[i] <- 0
        , df01$pop_z_20u[i] <- 1)
    )} 
    df01$pop_z_b10 <- 20
    for (i in 1:nrow(df01)){
    ifelse(
      ((as.integer(df01$pop_z_1020[i]) == as.integer(0))
       && (as.integer(df01$pop_z_20u[i]) == as.integer(0)))
      , df01$pop_z_b10[i] <- 1
      , ifelse(
        ((as.integer(df01$pop_z_1020[i]) == as.integer(1))
         || (as.integer(df01$pop_z_20u[i]) == as.integer(1)))
        , df01$pop_z_b10[i] <- 0
        , df01$pop_z_b10[i] <- 1)
    )}
 
```

# 2 - Matrix, removal of double 'edges', transposing, etc
```{r}

# create a distance matrix from two lists (-1 on both first and last city-row)
l1 <- data.frame(longitude = df01[1:387,3],
                 latitude = df01[1:387,4])
l2 <- data.frame(longitude = df01[2:388,3],
                 latitude = df01[2:388,4])
mtx <- distm(l1[,c('longitude','latitude')], l2[,c('longitude','latitude')], fun=distVincentyEllipsoid)

#dim(mtx)
# [1] 387 387
# elements in mtx = 149769
# checked: 387*387 = 149769

# remove duplicates and 'eigen'-distances
mtx[lower.tri(mtx)] <- NA
#head(mtx)
#         [,1]      [,2]     [,3]      [,4]      [,5]      [,6]
# [1,] 181411 158998.75 121147.8  45689.08 189958.68 204965.68
# [2,]     NA  60463.11 106807.8 178628.27  29788.73  46693.60
# [3,]     NA        NA 130832.1 141961.74  43773.50  51019.39
# [4,]     NA        NA       NA 145751.95 131649.72 150232.45
#dim(mtx) # > [1] 387 387

# transpose as quick-fix for correct iteration direction in as.list
mx <- t(mtx)
dists <- as.list(mx)
dists2 <- dists[!is.na(dists)]
dists3 <- data.frame(dists2)
dists4 <- t(dists3)
dists <- data.frame(dists4)

# ROUND_DISTANCES will be handy for further processing
# it introcuses a (tolerable?) error of <50 cm. (w/in one distance-measurement!)
dists$dists_km <- (dists$dists4 / 1000)
round_distances <- as.integer(round(dists$dists_km))
dists_z <- scale(round_distances, center = TRUE, scale = TRUE)

# dim(distances) # > [1] 75078     1 # checked: 149769 + 387 / 2 = 75078

```

# 3a - Get all 'pairwise' combinations; combine all variables
#    -- MERGING
```{r}
# dataframe should (row-wise) increase from 388 to 75078 (see above)
# get X-combinations per Y w/ 'combn' and split into seperate variables

# municipalities
compared_municipalities <- data.frame(combn(as.character(df01$municipality), 2, FUN = paste, collapse="_"))
# locations 
combined_lonlats <- paste(df01$longitude, df01$latitude, sep = ",")
compared_locations <- data.frame(combn(as.character(combined_lonlats), 2, FUN = paste, collapse="_"))
# populations

compared_pop_round <- data.frame(combn(as.character(df01$pop_round_100), 2, FUN = paste, collapse="_"))
compared_pop_z <- data.frame(combn(as.character(df01$pop_z), 2, FUN = paste, collapse="_"))
compared_pop_z_b10 <- data.frame(combn(as.character(df01$pop_z_b10), 2, FUN = paste, collapse="_"))
compared_pop_z_1020 <- data.frame(combn(as.character(df01$pop_z_1020), 2, FUN = paste, collapse="_"))
compared_pop_z_20u <- data.frame(combn(as.character(df01$pop_z_20u), 2, FUN = paste, collapse="_"))
 
df012 <- data.frame(compared_municipalities, compared_locations)
df013 <- data.frame(df012, dists)
df014 <- data.frame(df013, dists_z)
df015 <- data.frame(df014, compared_pop_round)
df016 <- data.frame(df015, compared_pop_z)
df017 <- data.frame(df016, compared_pop_z_b10)
df018 <- data.frame(df017, compared_pop_z_1020) 
df02 <- data.frame(df018, compared_pop_z_20u)

colnames(df02)[1] <- "compared_municipalities"
colnames(df02)[2] <- "compared_locations"
colnames(df02)[3] <- "distances"
colnames(df02)[4] <- "distances_round"
colnames(df02)[5] <- "dists_z_normalized"
colnames(df02)[6] <- "compared_pop_round"
colnames(df02)[7] <- "compared_pop_z"
colnames(df02)[8] <- "compared_pop_z_b10"
colnames(df02)[9] <- "compared_pop_z_1020"
colnames(df02)[10] <- "compared_pop_z_20u"
 
```

#    -- SPLITTING
```{r}
# make seperate lists with split-up's of compared variables

# munis
lcm <- data.frame(strsplit(as.character(df02$compared_municipalities), '_'))
lcm2 <- data.frame(t(lcm))
# locs
lcl <- data.frame(strsplit(as.character(df02$compared_locations), '_'))
lcl2 <- data.frame(t(lcl))
# posp
lcp <- data.frame(strsplit(as.character(df02$compared_pop_round), '_'))
lcp2 <- data.frame(t(lcp))
lcpz <- data.frame(strsplit(as.character(df02$compared_pop_z), '_'))
lcpz2 <- data.frame(t(lcpz))
lcpzb1 <- data.frame(strsplit(as.character(df02$compared_pop_z_b10), '_'))
lcpzb12 <- data.frame(t(lcpzb1))
lcp12 <- data.frame(strsplit(as.character(df02$compared_pop_z_1020), '_'))
lcp122 <- data.frame(t(lcp12))
lcp2u <- data.frame(strsplit(as.character(df02$compared_pop_z_20u), '_'))
lcp2u2 <- data.frame(t(lcp2u))

# merge back into one dataframe
df020 <- data.frame(df02, lcm2)
df021 <- data.frame(df020, lcl2)
df022 <- data.frame(df021, lcp2)
df023 <- data.frame(df022, lcpz2)
df024 <- data.frame(df023, lcpzb12)
df025 <- data.frame(df024, lcp122)
df03 <- data.frame(df025, lcp2u2)
